---
title: "Quantile Regression"
author: "Chris Cioffi, Kristina Frazier, Aidan Hennessy, Mike McHenry"
format: revealjs
editor: visual
---

## Quantile Regression

- A useful alternative to Ordinary Least Squares (OLS).

- Produces estimated regression coefficients for specific quantiles of the response variable.

  - As a result, the coefficients will change as the specified quantile of the response variable changes.
  
## Sample Density
  
```{r}
# Load libraries
library(tidyverse)

# Generate random normal data
df_rnd <- as.tibble(rnorm(10000))

df_rnd |>
  ggplot(aes(value)) +
  geom_density() +
  geom_vline(xintercept = quantile(df_rnd$value, prob = 0.25), color = "red") +
  geom_vline(xintercept = quantile(df_rnd$value, prob = 0.50), color = "black") +
  geom_vline(xintercept = quantile(df_rnd$value, prob = 0.99), color = "blue") +
  labs(title = "Sample Density Plot",
       x = "",
       y = "") +
  theme_bw()
```

  
## Why Use Quantile Regression?

- The relationship between the predictor and response variables may vary across the distribution, making it useful to allow the estimates to change.

- More robust to outliers.

- If OLS assumptions are violated.

  - Nonconstant variance, nonnormality, indpendence, nonlinearity.
  
## General Form of the Model

- $Q_\tau(Y) = \beta_0(\tau) + \beta_1(\tau)X_1 + ... + \beta_p(\tau)X_p$

  - Where $\tau$ is a specific quantile value.

- The objective of quantile regression is to minimize the median absolute deviations.

## Model Assumptions

## Quantreg

When you click the **Render** button a document will be generated that includes:

-   Content authored with markdown
-   Output from executable code

## References

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```
